---
title: "Montgomery County Crash Reporting Analysis"
author: "Lucy Murray"
date: "12/20/2022"
output: html_document
---

Source: [Open and Plot Shapefiles -- Introduction to Geospatial Raster and Vector Data with R (datacarpentry.org)](https://datacarpentry.org/r-raster-vector-geospatial/06-vector-open-shapefile-in-r/)

# Packages & Functions

## Load packages and supplemental functions.

```{r}
# Define packages to be used, install them if not already installed, and open them with library
packages <- c("readr", "dplyr", "sf", "terra", "spData", "tmap", "leaflet", "ggplot2", "purrr", "rgdal")

uninstalled_pkgs <- packages[!(packages %in% installed.packages()[, "Package"])]
if (length(uninstalled_pkgs) > 0) install.packages(uninstalled_pkgs)
invisible(lapply(packages, library, character.only = TRUE))

functions_to_source <- list.files("Functions/", full.names = TRUE)
walk(functions_to_source, source)
```

## Define Parameter Values

-   `user` unique to the system user.
-   `proj_dir` is the the local folder path; update as needed.

```{r}
user <- Sys.info()[["user"]]
proj_dir <- paste0("C:/Users/", user, "/Software/Git Local/mococrashreporting22/")
```

# Importing the Data

## 2022 Montgomery County Crashing Reporting data from [dataMontgomery](https://data.montgomerycountymd.gov/)

-   `062022_dataMontgomery_Crash_Reporting_-_Incidents_Data.csv` is the Comma Separated Value that contains the data being analyzed.

```{r}
mococrash2022 <- read_csv("062022_dataMontgomery_Crash_Reporting_-_Incidents_Data.csv")
```

## Montgomery County Zipcode Shapefile from [Montgomery County GIS Data](https://www.montgomerycountymd.gov/gis/data.html)

-   `Zipcodes_Montgomery` is the file name with the following extensions:
    -   `.shp` is the file that contains the geometry for all features.

    -   `.shx` is the file that indexes the geometry.

    -   `.dbf` is the file that stores feature attributes in a tabular form.

    -   `.prj` is the file that contains information on projection format including the coordinate system and projection information.

    -   `.sbn` and `.sbx`: are the files that are spatial indices of the features.

    -   `.shp.xml`: the file that is the geospatial metadata in XML format, (e.g. ISO 19115 or `XML` format).

The first shapefile that we will open contains the boundary of our study area (or our Area Of Interest or AOI, hence the name aoiBoundary_moco). To import shapefiles we use the sf function `st_read()`, which requires the file path to the shapefile.

```{r}
aoiBoundary_moco <- st_read(
  paste0(proj_dir, "Zipcodes_Montgomery.shp"))
```

Key metadata for all shapefiles include:

1.  **Object Type:** the class of the imported object.

2.  **Coordinate Reference System (CRS):** the projection of the data.

3.  **Extent:** the spatial extent (i.e. geographic area that the shapefile covers) of the shapefile. Note that the spatial extent for a shapefile represents the combined extent for all spatial objects in the shapefile.

We can view shapefile metadata using the `st_geometry_type()`, `st_crs()` and `st_bbox()` functions. First, let's view the geometry type for our AOI shapefile:

```{r}
st_geometry_type(aoiBoundary_moco)
```

Our `aoi_boundary_HARV` is a polygon object. The 18 levels shown below our output list the possible categories of the geometry type. Now let's check what CRS this file data is in:

```{r}
st_crs(aoiBoundary_moco)
```

Our data in the CRS **Maryland (ftUS)**. The CRS is critical to interpreting the object's extent values as it specifies units. To find the extent of our AOI, we can use the `st_bbox()` function:

```{r}
st_bbox(aoiBoundary_moco)
```

# Plotting

## Creating the city labels

```{r}
moco_points <- st_centroid(aoiBoundary_moco)
moco_points <- cbind(aoiBoundary_moco, st_coordinates(st_centroid(aoiBoundary_moco$geometry)))
```

## Shape boundary only

```{r}
ggplot() + 
  geom_sf(data = aoiBoundary_moco, aes(fill=CITY), show.legend = FALSE) + 
  geom_text(data = moco_points, aes(x=X, y=Y, label=CITY), color = "darkblue", fontface = "bold", check_overlap = TRUE) +
  ggtitle("Montgomery County Boundary Plot", subtitle = paste0("(", length(unique(aoiBoundary_moco$CITY)), " Cities)")) + 
  xlab("Latitude") +
  ylab("Longitude") + 
  coord_sf()
```

Plotting the data

```{r}
crash_locations <- mococrash2022 %>%
  filter(`Hit/Run`=="Yes") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(clusterOptions = markerClusterOptions(lng = ~Longitude, lat = ~Latitude, popup = ~`Report Number`, label = ~`Report Number`))
crash_locations
```
